---
title: "Opdracht1.2"
author: "Laurine Seelt"
date: '2022-04-21'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(readxl)
```

https://pubmed.ncbi.nlm.nih.gov/29324233/

Study purpose: 1

Data availability statement: 1

Data location: 1

Study location: 1

Author review: 7

Ethics statement: 0

Funding statement: 1

Code availability: 1 ?


In this study, researchers investigated the heterologous effects of Bacillus Calmette-Guérin (BCG) vaccination against other, non-related virus infections. 
It has been known for some time that BCG vaccination has positive heterologous effects against some other virusses, but the exact immunological basis for this effect has never been examined in depth. The researchers in this study examined the genome-wide histone modifications that take place as a result of BCG vaccination, and also examined the functional changes in monocytes that take place after BCG vaccination. The effect of BCG vaccination on viral, serological and immunological parameters after administration of the Yellow Fever Vaccine was also examined.

The method used for this study was ChIP-sequencing, to look at genome-wide changes in the distribution of H3K27ac. H3K27ac is a marker of active promoters and enhancers. This was done before BCG vaccination and one month after BCG vaccination. Correlation plots showed a clear distinction between the two kinds of samples. Most changes between the two kinds of samples were related to G-protein coupled receptors and protein kinases, which shows that BCG vaccination has a substantial effect on remodeling signal transduction molecules.


__Part 2__

https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0252047

The code tries to achieve ..

The readibility of the code is very good, 5.

__Trying to recreate figure 2.__

```{r}
library(tidyverse)
library(here)
library(reticulate)
# py_install("numpy")
# py_install("pandas")
# py_install(packages = "matplotlib")
# py_install("seaborn")
# py_install(packages = "scipy")
```


```{python, eval = FALSE}
#Import the packages needed to make this notebook work

import numpy as np
import pandas as pd
import matplotlib.pyplot as plt
import seaborn as sns
import scipy.stats as stats

#Make sure the figures appear in the notebook

%matplotlib inline

#Import the data

df = pd.read_csv("Data_raw/Borghi_VanGulick_PsychRDM_Data.csv")
```

