---
author: "Laurine Seelt"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
biblio-style: apalike
link-citations: yes
---

# Cleaning up according to the Guerrilla analytics framework

This was the third assignment for our portfolio. We had to apply the Guerrilla analytics framework to our DAUR2 folder.
<br>
I decided to: <br>
1. Make a folder called "Rmd" in which I keep all of my markdowns. <br>
2. Make a folder called "Data" in which I store all the data I've used. This folder is split in "Metagenomics" and "RNA_seq", to separate the two anlyses I've conducted during Daur2. Both those folders contain the sub folder "Data_raw", which is data I cannot touch. This is the raw data I received from teachers or downloaded from the internet. <br>
3. Make a folder called "R". In this folder, I store functions I've written. <br>
4. Make a folder called "Code". This folder contains the scripts I've written, except my Rmarkdowns.
<br>
```{r setup3, include=FALSE}
library(tidyverse)
# install.packages("fs")
library(fs)
library(here)
```

```{r}
fs::dir_tree(path = here("Daur2"), type = "directory")
```

