---
author: "Laurine Seelt"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
biblio-style: apalike
link-citations: yes
---

# Working with DBeaver and SQL

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(dslabs)
library(here)
library(DBI)
# install.packages("RPostgres")
library(RPostgres)
```

First, I read in the flu, dengue and gapminder data.
<br>
```{r, message=FALSE, warning=FALSE}
flu <- read_csv(here("data/data_raw/flu_data.csv"), skip = 11)
dengue <- read_csv(here("data/data_raw/dengue_data.csv"), skip = 11)
gapminder <- read_builtin("gapminder")
```

The flu and dengue data wasn't tidy yet, so I solved that by using the pivot_longer command to create the column "country" and "cases" in each of the datasets. That way, each country, date and number of cases had its own column. <br>
The gapminder dataset was already tidy.
<br>
```{r making the data tidy, message=FALSE, warning=FALSE}
dengue_tidy <- pivot_longer(data = dengue, cols = c("Argentina", "Bolivia", "Brazil", "India", "Indonesia", "Mexico", "Philippines", "Singapore", "Thailand", "Venezuela"), names_to = "country_d", values_to = "dengue_cases")
flu_tidy <- pivot_longer(data = flu, cols = c("Argentina", "Australia", "Austria", "Belgium", "Bolivia", "Brazil", "Bulgaria", "Canada", "Chile", "France", "Germany", "Hungary", "Japan", "Mexico", "Netherlands", "New Zealand", "Norway", "Paraguay", "Peru", "Poland", "Romania", "Russia", "South Africa", "Spain", "Sweden", "Switzerland", "Ukraine", "United States", "Uruguay"), names_to = "country_f", values_to = "flu_cases")
```
<br>
Next, I changed the flu and dengue data some more. I separated the Data column into the three columns "Year", "Month" and "Day". I also added a d or f after the column name "year", so joining of the datasets in DBeaver will be possible later on.
In the gapminder dataset, I changed the column "year" to "year_g" to also make joining more easy later on in DBeaver.
<br>
```{r some data wrangling, message=FALSE, warning=FALSE}
flu_tidy <- separate(flu_tidy, Date, into = c("year_f", "month", "day"), sep = "-")
dengue_tidy<- separate(dengue_tidy, Date, into = c("year_d", "month", "day"), sep = "-")
flu_tidy$country_flu <- as.factor(flu_tidy$country_f)
flu_tidy$year <- as.numeric(flu_tidy$year_f)
dengue_tidy$country_dengue <- as.factor(dengue_tidy$country_d)
dengue_tidy$year <- as.numeric(dengue_tidy$year_d)
gapminder$country <- as.factor(gapminder$country)
dengue_tidy <- as.data.frame(dengue_tidy)
flu_tidy <- as.data.frame(flu_tidy)
dengue_tidy <- dengue_tidy %>% group_by(country_d, year_d) %>% summarise_each(funs(sum), dengue_cases)
flu_tidy <- flu_tidy %>% group_by(country_f, year_f) %>% summarise_each(funs(sum), flu_cases)
# Thanks to https://stackoverflow.com/questions/25089665/error-only-defined-on-a-data-frame-with-all-numeric-variables-with-ddply-on-lar for the summarise_each command :)
gapminder <- gapminder %>% rename(year_g = "year", country_g = "country")
gapminder$year_g <- as.character(gapminder$year_g)
```
<br>
Then I saved the three datasets as both .csv files and .rds files.
<br>
```{r saving .csv and .rds files, message=FALSE, warning=FALSE}
write.csv(flu_tidy, "Data//flu_data_tidy.csv", row.names = TRUE)
write.csv(dengue_tidy, "Data//dengue_data_tidy.csv", row.names = TRUE)
write.csv(gapminder, "Data//gapminder.csv", row.names = TRUE)
saveRDS(flu_tidy, file = "flu_data_tidy.rds")
saveRDS(dengue_tidy, file = "dengue_data_tidy.rds")
saveRDS(gapminder, file = "gapminder.rds")
```
<br>
Next, I inserted the tables into the "workflowsdb" database in DBeaver.
<br>
```{r exporting the files, message=FALSE, warning=FALSE, eval=FALSE}
con <- dbConnect(RPostgres::Postgres(),
                 dbname = "workflowsdb",
                 host = "localhost",
                 port = "5432",
                 user = "postgres",
                 password = "SQL22")
dbWriteTable(con, "flu_tidy", flu_tidy)
dbWriteTable(con, "dengue_tidy", dengue_tidy)
dbWriteTable(con, "gapminder", gapminder)
dbDisconnect(con)
```
<br>
Then I inspected the contents of the datasets in DBeaver and in R.
<br>
```{r inspecting the contents in R, message=FALSE, warning=FALSE}
# Inspecting the flu_tidy dataset
flu_tidy %>% arrange(desc(flu_cases)) %>% head()
flu_tidy$flu_cases %>% min(flu_tidy$flu_cases, na.rm = TRUE)
flu_tidy$flu_cases %>% max(flu_tidy$flu_cases, na.rm = TRUE)
# Inspecting the dengue_tidy dataset
dengue_tidy %>% arrange(desc(dengue_cases)) %>% head()
dengue_tidy$dengue_cases %>% min(dengue_tidy$dengue_cases, na.rm = TRUE)
dengue_tidy$dengue_cases %>% max(dengue_tidy$dengue_cases, na.rm = TRUE)
# Inspecting the gapminder dataset
gapminder$infant_mortality %>% min(gapminder$infant_mortality, na.rm = TRUE)
gapminder$infant_mortality %>% max(gapminder$infant_mortality, na.rm = TRUE)
gapminder$life_expectancy %>% min(gapminder$life_expectancy, na.rm = TRUE)
gapminder$life_expectancy %>% max(gapminder$life_expectancy, na.rm = TRUE)
gapminder %>% select(country_g, year_g, life_expectancy) %>% arrange(desc(life_expectancy)) %>% head
gapminder %>% select( country_g, year_g, infant_mortality) %>% arrange(desc(infant_mortality)) %>% head()
```
<br>
