---
author: "Laurine Seelt"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
biblio-style: apalike
link-citations: yes
---

# SQL

To show that I can work with SQL in DBeaver, I worked with three data sets that I inspected, joined and I made some figures with the joined data set.
<br>
```{r, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(readxl)
library(dslabs)
library(here)
library(DBI)
# install.packages("RPostgres")
library(RPostgres)
```
<br>
First, I read in the flu, dengue and gapminder data.
<br>
```{r reading in the three files, message=FALSE, warning=FALSE}
flu <- read_csv(here("data/data_raw/flu_data.csv"), skip = 11)
dengue <- read_csv(here("data/data_raw/dengue_data.csv"), skip = 11)
gapminder <- read_builtin("gapminder")
```
<br>
The flu and dengue data wasn't tidy yet, so I solved that by using the pivot_longer command to create the column "country" and "cases" in each of the datasets. That way, each country, date and number of cases had its own column. <br>
The gapminder data set was already tidy.
<br>
```{r making the data tidy, message=FALSE, warning=FALSE}
dengue_tidy <- pivot_longer(data = dengue, cols = c("Argentina", "Bolivia", "Brazil", "India", "Indonesia", "Mexico", "Philippines", "Singapore", "Thailand", "Venezuela"), names_to = "country_d", values_to = "dengue_cases")
flu_tidy <- pivot_longer(data = flu, cols = c("Argentina", "Australia", "Austria", "Belgium", "Bolivia", "Brazil", "Bulgaria", "Canada", "Chile", "France", "Germany", "Hungary", "Japan", "Mexico", "Netherlands", "New Zealand", "Norway", "Paraguay", "Peru", "Poland", "Romania", "Russia", "South Africa", "Spain", "Sweden", "Switzerland", "Ukraine", "United States", "Uruguay"), names_to = "country_f", values_to = "flu_cases")
```
<br>
Next, I changed the flu and dengue data some more. I separated the Data column into the three columns "Year", "Month" and "Day". I also added a 'd' or 'f' after the column name "year", so joining of the data sets in DBeaver will be possible and easy later on.
In the gapminder data set, I changed the column "year" to "year_g" to also make joining easier later on in DBeaver.
<br>
```{r some data wrangling, message=FALSE, warning=FALSE}
flu_tidy <- separate(flu_tidy, Date, into = c("year_f", "month", "day"), sep = "-")
dengue_tidy<- separate(dengue_tidy, Date, into = c("year_d", "month", "day"), sep = "-")
flu_tidy$country_flu <- as.factor(flu_tidy$country_f)
flu_tidy$year <- as.numeric(flu_tidy$year_f)
dengue_tidy$country_dengue <- as.factor(dengue_tidy$country_d)
dengue_tidy$year <- as.numeric(dengue_tidy$year_d)
gapminder$country <- as.factor(gapminder$country)
dengue_tidy <- as.data.frame(dengue_tidy)
flu_tidy <- as.data.frame(flu_tidy)
dengue_tidy <- dengue_tidy %>% group_by(country_d, year_d) %>% summarise_each(funs(sum), dengue_cases)
flu_tidy <- flu_tidy %>% group_by(country_f, year_f) %>% summarise_each(funs(sum), flu_cases)
# Thanks to https://stackoverflow.com/questions/25089665/error-only-defined-on-a-data-frame-with-all-numeric-variables-with-ddply-on-lar for the summarise_each command :)
gapminder <- gapminder %>% rename(year_g = "year", country_g = "country")
gapminder$year_g <- as.character(gapminder$year_g)
```
<br>
Then I saved the three datasets as both .csv files and .rds files.
<br>
```{r saving .csv and .rds files, message=FALSE, warning=FALSE, eval=FALSE}
write.csv(flu_tidy, "Data//flu_data_tidy.csv", row.names = TRUE)
write.csv(dengue_tidy, "Data//dengue_data_tidy.csv", row.names = TRUE)
write.csv(gapminder, "Data//gapminder.csv", row.names = TRUE)
saveRDS(flu_tidy, file = "flu_data_tidy.rds")
saveRDS(dengue_tidy, file = "dengue_data_tidy.rds")
saveRDS(gapminder, file = "gapminder.rds")
```
<br>
Next, I inserted the tables into the "workflowsdb" database in DBeaver that I had already made.
<br>
```{r exporting the files, message=FALSE, warning=FALSE, eval=FALSE}
con <- dbConnect(RPostgres::Postgres(),
                 dbname = "workflowsdb",
                 host = "localhost",
                 port = "5432",
                 user = "postgres",
                 password = "")
dbWriteTable(con, "flu_tidy", flu_tidy)
dbWriteTable(con, "dengue_tidy", dengue_tidy)
dbWriteTable(con, "gapminder", gapminder)
dbDisconnect(con)
```
<br>
Then I inspected the contents of the datasets in DBeaver, with the following code:
<br>
<br>
select dengue_cases, country <br>
from dengue_tidy <br>
order by dengue_cases asc <br>
<br>
select dengue_cases, country <br>
from dengue_tidy <br>
order by dengue_cases desc <br>
<br>
select flu_cases, country <br>
from flu_tidy <br>
order by flu_cases asc <br>
<br>
select flu_cases, country <br>
from flu_tidy <br>
order by flu_cases desc <br>
<br>
select infant_mortality, country <br>
from gapminder <br>
order by infant_mortality desc <br>
<br>
select infant_mortality, country <br>
from gapminder <br>
order by infant_mortality asc <br>
<br>
select life_expectancy, country <br>
from gapminder <br>
order by life_expectancy desc <br>
<br>
select life_expectancy, country <br>
from gapminder <br>
order by life_expectancy asc
<br>
<br>
I also inspected the data in R, with the code below:
<br>
```{r inspecting the contents in R, message=FALSE, warning=FALSE}
## Inspecting the flu_tidy dataset
flu_tidy %>% arrange(desc(flu_cases)) %>% head()
flu_tidy$flu_cases %>% min(flu_tidy$flu_cases, na.rm = TRUE)
flu_tidy$flu_cases %>% max(flu_tidy$flu_cases, na.rm = TRUE)
flu_tidy$country_f %>% unique()

## Inspecting the dengue_tidy dataset
dengue_tidy %>% arrange(desc(dengue_cases)) %>% head()
dengue_tidy$dengue_cases %>% min(dengue_tidy$dengue_cases, na.rm = TRUE)
dengue_tidy$dengue_cases %>% max(dengue_tidy$dengue_cases, na.rm = TRUE)
dengue_tidy$country_d %>% unique()

## Inspecting the gapminder dataset
gapminder$infant_mortality %>% min(gapminder$infant_mortality, na.rm = TRUE)
gapminder$infant_mortality %>% max(gapminder$infant_mortality, na.rm = TRUE)
gapminder$life_expectancy %>% min(gapminder$life_expectancy, na.rm = TRUE)
gapminder$life_expectancy %>% max(gapminder$life_expectancy, na.rm = TRUE)
gapminder %>% select(country_g, year_g, life_expectancy) %>% arrange(desc(life_expectancy)) %>% head
gapminder %>% select( country_g, year_g, infant_mortality) %>% arrange(desc(infant_mortality)) %>% head()
gapminder$country_g %>% unique() %>% head(20)
```
<br>
In DBeaver, I performed some joins and exported a joined table to R. The joined table is called "Joined_table.csv".
<br>

The SQL code that I used to create the joined table is as follows:

select *
from flu_tidy
left join dengue_tidy on year_f = year_d
and country_f = country_d
left join gapminder on year_f = year_g
and country_f = country_g

<br>
After reading in the data from "Joined_table.csv", I removed the rows "year_d", "year_g", "country_d" and "country_g". These rows were redundant, seeing as we already have the rows "year_f" and "country_f" from the flu dataset. I also changed the name of "country_f" to "country" and "year_f" to "year".
<br>
```{r generating a tidy joined table, message=FALSE, warning=FALSE}
joined_table <- read.csv(here("Data/Joined_table.csv"))

joined_table <- joined_table %>% select(country_f:region, -(country_d:year_d))
joined_table <- joined_table %>% select(country_f:region, -(country_g:year_g))
joined_table <- joined_table %>% rename(country = "country_f", year = "year_f")
```
<br>
Then I performed some descriptive statistics with this data set and made some visualizations.
<br>
```{r descriptive statistics, message=FALSE, warning=FALSE}
joined_table %>% summary()
```
<br>
For the graphs from joined_table, I only selected the countries that were relevant. So, for the dengue graph, I only selected the countries Argentina, Bolivia, Brazil, India, Indonesia, Mexico, Philippines, Singapore, Thailand and Venezuela because those were the only countries present in the original dengue data set, as we saw earlier. <br>
<br>
```{r dengue graph, message=FALSE, warning=FALSE}
joined_table %>% group_by(country) %>% filter(country == "Argentina" | country == "Bolivia" | country == "Brazil" | country == "India" | country == "Indonesia" | country == "Mexico" | country == "Philippines" | country == "Singapore" | country == "Thailand" | country == "Venezuela") %>%
  ggplot(aes(x = year, y = dengue_cases, colour = country)) +
  geom_point() +
  geom_line() +
  theme_classic() +
  scale_x_continuous(breaks = seq(2002, 2015, 1))+
  labs(title = "Dengue cases in the years 2002 - 2015",
       x = "Year",
       y = "Dengue cases")
```
<br>
In the dengue graph, only the countries Argentina, Bolivia, Brazil and Mexico are present, eventhough I selected more countries. Maybe the other countries aren't present in our joined_table anymore. Let's take a look at the countries in joined_table.
<br>
```{r inspecting the countries, message=FALSE, warning=FALSE}
joined_table$country %>% unique()
```
<br>
As expected, India, Indonesia, Philippines, Singapore, Thailand and Venezuela aren't present in joined_table. <br>
Let's see if these countries also have data about flu incidence.
<br>
```{r flu graph, message=FALSE, warning=FALSE}
joined_table %>% group_by(country) %>% filter(country == "Argentina" | country == "Bolivia" | country == "Brazil"| country == "Mexico") %>%
  ggplot(aes(x = year, y = flu_cases, colour = country)) +
  geom_point() +
  geom_line() +
  theme_classic() +
  scale_x_continuous(breaks = seq(2002, 2015, 1))+
  labs(title = "Flu cases in the years 2002 - 2015",
       x  = "Year", 
       y = "Flu cases")
```
<br>
Lastly, let's see if these countries also have data about infant mortality, life expectancy, fertility and population.
<br>
```{r gapminder graphs, message=FALSE, warning=FALSE}
joined_table %>% group_by(country) %>% filter(country == "Argentina" | country == "Bolivia" | country == "Brazil"| country == "Mexico") %>%
  ggplot(aes(x = year, y = infant_mortality, colour = country)) +
  geom_point() +
  geom_line() +
  theme_classic() +
  scale_x_continuous(breaks = seq(2002, 2015, 1)) +
  labs(title = "Infant mortality in the years 2002 - 2015",
       x = "Year",
       y = "Infant mortality")

joined_table %>% group_by(country) %>% filter(country == "Argentina" | country == "Bolivia" | country == "Brazil" | country == "Mexico") %>%
  ggplot(aes(x = year, y = life_expectancy, colour = country)) +
  geom_point() +
  geom_line() +
  theme_classic() +
  scale_x_continuous(breaks = seq(2002, 2015, 1)) +
  labs(title = "Life expectancy in the years 2002 - 2015",
       x = "Year",
       y = "Life expectancy")

joined_table %>% group_by(country) %>% filter(country == "Argentina" | country == "Bolivia" | country == "Brazil" | country == "Mexico") %>%
  ggplot(aes(x = year, y = fertility, colour = country)) +
  geom_point() +
  geom_line() +
  theme_classic() +
  scale_x_continuous(breaks = seq(2002, 2015, 1)) +
  labs(title = "Fertility in the years 2002 - 2015",
       x = "Year",
       y = "Fertility")

joined_table %>% group_by(country) %>% filter(country == "Argentina" | country == "Bolivia" | country == "Brazil" | country == "Mexico") %>%
  ggplot(aes(x = year, y = population, colour = country)) +
  geom_point() +
  geom_line() +
  theme_classic() +
  scale_x_continuous(breaks = seq(2002, 2015, 1)) +
  labs(title = "Population in the years 2002 - 2015",
       x = "Year",
       y = "Population")
```
