---
author: "Laurine Seelt"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
biblio-style: apalike
link-citations: yes
---

# Package

Packages can be very useful when working in R. Here, I demonstrate how I've created the package called covergirl. <br>
<br>
```{r setting up, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(dragracer)
library(covergirl)
```
This package can be used to search for some information about contestants of the show RuPaul's Drag Race.
<br>
The package can be installed by running devtools::install_github("laurineseelt/covergirl", build_vignettes = TRUE). The vignette can be found by running vignette("covergirl") or browseVignettes("covergirl").
<br>
<br>
With the function general_info(), you can type in the name of a contestant and you'll see the season that queen competed on, their age when they first competed on RuPaul's Drag Race and their hometown.
<br>
```{r voorbeeld general_info, message=FALSE, warning=FALSE}
general_info("Jinkx Monsoon")
```
<br>
With the function episode_name(), you can type in the number of the season and the episode, and this function will tell you the name of that particular episode.
<br>
```{r example of episode_name, message=FALSE, warning=FALSE}
episode_name("S10", 5)
```
<br>
With the function special_ep(), you can type in the number of the season and the episode, and it'll tell you if this was a special episode of RuPaul's Drag Race or not.
<br>
```{r example special_ep, message=FALSE, warning=FALSE}
special_ep("S14", 11)
special_ep("S13", 5)
```
<br>
Lastly, with the function winner(), you can type in the number of a season and this function which queen won that particular season.
<br>
```{r example winner, message=FALSE, warning=FALSE}
winner("S05")
```
<br>
The vignette for this package can be found 