---
author: "Laurine Seelt"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
params:
  country: 
    value: Spain
  year:
    value: 2022
  month: 
    value: February
documentclass: book
biblio-style: apalike
link-citations: yes
---

# Parameterization

To show that I can work with parameterization in my Markdowns, I made a parameterized markdown with data about COVID-19 cases and COVID-19 related deaths.
```{r loading the packages, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(here)
library(ggplot2)
```
<br>
The data used for this markdown was from the ECDC website. To start, I downloaded this data from the website and read it in in RStudio.
<br>
```{r reading in the covid-19 data, message=FALSE, warning=FALSE}
covid19_data <- read_csv(here::here("Data/Data_raw/covid19_data.csv"))
```
<br>
Then, I stored the three parameters as objects so I can use them in the title of my figures. I also changed the value 'month' to a numeric variable, so it'd be easier to use later on in the graphs.
<br>
```{r making an object of the country, message=FALSE, warning=FALSE, eval=FALSE}
covid19_data <- covid19_data %>% mutate(month = month.name[as.numeric(month)])
country <- params$country
year <- params$year
months <- params$month
# covid19_data_filtered <- covid19_data %>% filter(year %in% params$year, month %in% params$month, countriesAndTerritories %in% params$country)
```
<br>
I created a graph to show the COVID-19 cases per day for a specific month, and I also created a graph for the amount of COVID-19 related deaths per day for a specific month.
<br>
```{r visualising the covid-19 cases, message=FALSE, warning=FALSE, eval=FALSE}
covid19_data %>% group_by(day, month, year) %>% filter(countriesAndTerritories == params$country & year == params$year & month == params$month) %>%
  ggplot(aes(x = day, y = cases)) +
  geom_point() +
  scale_x_continuous(breaks = seq(1, 31, 1)) +
  labs(title = paste0("COVID-19 cases in ", country, ", ", months, " ", year),
       x = "Day",
       y = "COVID-19 cases")
```

```{r visualising the covid-19 related deaths, message=FALSE, warning=FALSE, eval=FALSE}
covid19_data %>% group_by(day, month, year) %>% filter(countriesAndTerritories == params$country & year == params$year & month == params$month) %>%
  ggplot(aes(x = day, y = deaths)) +
  geom_point() +
  scale_x_continuous(breaks = seq(1, 31, 1)) +
  labs(title = paste0("COVID-19 related deaths in ", country, ", ", months, " ", year),
       x = "Day" ,
       y = "COVID-19 related deaths")
```

I made three HTML's with rmarkdown::render() in R.

<div>

<a> href='parameterization_ireland.html#Parameterization_Ireland'>For a parameterized HTML on the COVID-19 cases and deaths in Ireland, August 2020, click here<a/>

<a> href='parameterization_belgium.html#Parameterization_Belgium'>For a parameterized HTML on the COVID-19 cases and deaths in Belgium, April 2021, click here<a/>

<a> href="parameterization_austria.html#Parameterization_Austria'>For a parameterized HTML on the COVID-19 cases and deaths in Austria, February 2022, click here<a/>

</div>