---
author: "Laurine Seelt"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
params:
  country: 
    value: Spain
  year: 
    value: 2021
  month: 
    value: April
documentclass: book
biblio-style: apalike
link-citations: yes
---

# Spain Parameters

To show that I can work with parameterization in my Markdowns, I made a parameterized markdown with data about COVID-19 cases and COVID-19 related deaths.
```{r loading the packages2, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(here)
library(ggplot2)
```
<br>
The data used for this markdown was from the ECDC website. To start, I downloaded this data from the website and read it in in RStudio.
<br>
```{r reading in the covid-19 data2, message=FALSE, warning=FALSE}
covid19_data <- read_csv(here::here("Data/Data_raw/covid19_data.csv"))
```
<br>
Then, I stored the three parameters as objects so I can use them in the title of my figures. I also changed the value 'month' to a numeric variable, so it'd be easier to use later on in the graphs.
<br>
```{r making an object of the country2, message=FALSE, warning=FALSE}
covid19_data <- covid19_data %>% mutate(month = month.name[as.numeric(month)])
#country <- params$country
#year <- params$year
#months <- params$month
parameters <- list()
parameters$country <- "Spain"
parameters$year <- 2021
parameters$month <- "April"
covid19_data_filtered <- covid19_data %>% filter(year %in% parameters$year, 
                                                 countriesAndTerritories %in% parameters$country, 
                                                 month %in% parameters$month)
```
<br>
I created a graph to show the COVID-19 cases per day for a specific month, and I also created a graph for the amount of COVID-19 related deaths per day for a specific month. In this markdown, I did this for Belgium.
<br>
```{r visualising the covid-19 cases2, message=FALSE, warning=FALSE}
covid19_data %>% group_by(day, month, year) %>% 
  filter(countriesAndTerritories == parameters$country & 
         year == parameters$year & 
         month == parameters$month) %>%
  ggplot(aes(x = day, y = cases)) +
  geom_point() +
  scale_x_continuous(breaks = seq(1, 31, 1)) +
  labs(title = paste0("COVID-19 cases in ", parameters$country, ", ", parameters$month, " ", parameters$year),
       x = "Day",
       y = "COVID-19 cases")
```

```{r visualising the covid-19 related deaths2, message=FALSE, warning=FALSE}
covid19_data %>% group_by(day, month, year) %>% 
  filter(countriesAndTerritories == parameters$country & 
         year == parameters$year & 
         month == parameters$month) %>%
  ggplot(aes(x = day, y = deaths)) +
  geom_point() +
  scale_x_continuous(breaks = seq(1, 31, 1)) +
  labs(title = paste0("COVID-19 related deaths in ", parameters$country, ", ", parameters$month, " ", parameters$year),
       x = "Day" ,
       y = "COVID-19 related deaths")
```

<div>
<a href='ireland-parameters.html#Ireland Parameters'>For a parameterized HTML on the COVID-19 cases and deaths in Ireland, August 2020, click here<a/>

<a href='austria-parameters.html#Austria Parameters'>For a parameterized HTML on the COVID-19 cases and deaths in Austria, February 2022, click here<a/>

</div>